---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(matos)
```


## List available files

First, `list_projects` returns the [project page](https://matos.asascience.com/project), which is useful to figure out what URLs are associated with each project. You do not need MATOS permissions in order to view this page.

```{r list_projs}
all_projects <- list_projects()

head(all_projects)
```

I can also view the files that I've uploaded to my projects using `list_project_files`, but that requires logging in first. The family of `list_` functions in this package will prompt you to log in before moving on. Note that I'll be entering my MATOS username and password behind the scenes here.

```{r}
project_files <- list_project_files(project = 'umces boem offshore wind energy')

head(project_files)
```

### User credentials
A side note on your MATOS username and password: `matos` defaults to asking you for your login credentials every time you start a new session. To skirt around this you can use `set_matos_credentials()`, which installs your username and password in your [.Renviron file](https://rstats.wtf/r-startup.html#renviron). You will be automatically logged in every time you use your current computer after doing this, but beware: someone else could theoretically access your username and password if they gain access to your computer.

### Back to the regularly-scheduled programming
I can also list any of my OTN node *Data Extraction Files*.

```{r}
ACT_MATOS_files <- list_extract_files(project = 'umces boem offshore wind energy',
                                      detection_type = 'all')

head(ACT_MATOS_files)
```


## Download project or data extraction files

There are a few ways to download the different types of files held by MATOS. I can download directly if I know the URL of the file:

```{r}
project_files$url[1]

get_project_file(url = project_files$url[1])
```

I can download by using an index from the `ACT_MATOS_files` table above, here the file on the second row.

```{r}
get_extract_file(file = 2, project = 'umces boem offshore wind energy')
```

## Search and download tag detections

Using the `tag_search` function, I can interface with MATOS' [tag search page](https://matos.asascience.com/search). Be very careful with this function -- it can take a *very*, **VERY** long time to return your files. This function downloads the requested CSV into your working directory, and, if `import = T` is used, reads it into your R session.

```{r, eval=FALSE}
my_detections <- tag_search(tags = paste0('A69-1601-254', seq(60, 90, 1)),
                            start_date = '03/01/2016',
                            end_date = '04/01/2016', 
                            import = T)
```

## Upload files to MATOS

There are times when you want to upload new data to MATOS. The currently accepted data types and formats are:

- newly-deployed transmitters (CSV/XLS(X))
- detection logs (CSV/VRL)
- receiver and glider deployment metadata (CSV/XLS(X))
- receiver events like temperature, tilt, etc. (CSV)
- glider GPS tracks (CSV)

A few data types use designated Ocean Tracking Network templates:

- tag metadata
- receiver deployment metadata
- glider deployment metadata

If you don't have one of these templates downloaded, you can download it through the package. For example:
```{r}
get_otn_template('glider')
```

Then, get to uploading!
```{r eval=FALSE}
upload_file(project = 'umces boem offshore wind energy',
            file = c('this_is_a_dummy_file.csv', 'so_is_this.csv'),
            data_type = 'new_tags')
```

