---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# matos

<!-- badges: start -->
<!-- badges: end -->

{matos} is a wrapper over a suite of [httr](https://httr.r-lib.org/) and [rvest](https://rvest.tidyverse.org/) functions made to interact with the [Mid-Atlantic Acoustic Telemetry Observing System website](https://matos.asascience.com/). Because of this, it's not necessarily fast or the best way to do things -- we're just pinging the website back and forth. Additionally, an internet connection is needed for pretty much anything to work.

Please note that you will need a MATOS account, [which you can sign up for here](https://matos.asascience.com/account/signup), in order to interface with any project-specific files.

## Installation

This package is still in "throw things on the wall and see what sticks" phase. You can install the most-up-to-date version from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("mhpob/matos")
```
## Quick start

The first thing you should do is log into your MATOS account. This can be done through the package, using the [RStudio API](https://rstudio.github.io/rstudio-extensions/rstudioapi.html). This probably won't work if you're not using RStudio, so it will be changed in the future. A pop up will appear asking for your username and password. If everything works out, your credentials will be kept in the sessions' cookies. Your username/password will not be saved -- this was done intentionally so that you don't accidentally save credentials in a public script.

```{r example}
library(matos)

matos_login()
```


In addition to the above, there are/will be a few fucntions that take no arguments -- their sole function is to ping the website and return the data to you. `matos_projects` returns the [project page](https://matos.asascience.com/project), which is useful to figure out what URLs are associated with each project. You do not need MATOS permissions in order to view this page.

```{r list_projs}
all_projects <- matos_projects()

head(all_projects)
```

Since I'm logged in, I can view the files that I've uploaded to my projects.

```{r}
files <- project_files(project = 'umces boem offshore wind energy')

head(files)
```

And, I can download any of my choosing.

```{r}
files$url[1]

get_project_file(file = files$url[1])
```

