---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit this file -->
<!-- Very likely that you'll need to run rmarkdown::render('readme.rmd') rather than the knit button. -->
<!-- readme.html will be created and is unnecessary, so delete that. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# matos

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/mhpob/matos/workflows/R-CMD-check/badge.svg)](https://github.com/mhpob/matos/actions)
<!-- badges: end -->

{matos} is an attempt at an API to the [Mid-Atlantic Acoustic Telemetry Observing System website](https://matos.asascience.com/), powered by a suite of [httr](https://httr.r-lib.org/) and [rvest](https://rvest.tidyverse.org/) functions. Because of this, it's not necessarily fast or the best way to do things -- we're just pinging the website back and forth. Additionally, an internet connection is needed for pretty much anything to work.

Please note that you will need a MATOS account, [which you can sign up for here](https://matos.asascience.com/account/signup), in order to interface with any project-specific files.

## News!
### 2021-04-12
  - `get_updates`: A new function to download all files updated since a given date. Super useful after a data push!
  - `list_files` now has a `since` argument, allowing you to only list the files that have been updated since a certain date.

## Installation

This package is still in "throw things on the wall and see what sticks" phase (*"this package is currently undergoing heavy development"*). You can install the most-up-to-date version from GitHub:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("mhpob/matos")
```

## List available files

First, `list_projects` returns the [project page](https://matos.asascience.com/project), which is useful to figure out what URLs are associated with each project. You do not need MATOS permissions in order to view this page.

```{r list_projs}
library(matos)

all_projects <- list_projects()

head(all_projects)
```

I can also view the files that I've uploaded to my projects using `list_project_files`, but that requires logging in first.

```{r}
project_files <- list_project_files(project = 'umces boem offshore wind energy')

head(project_files)
```

I can also list any of my OTN node *Data Extraction Files*.

```{r}
ACT_MATOS_files <- list_extract_files(project = 'umces boem offshore wind energy', detection_type = 'all')

head(ACT_MATOS_files)
```


## Download project or data extraction files

There are a few ways to download the different types of files held by MATOS. I can download directly if I know the URL of the file:

```{r}
project_files$url[1]

get_file(url = project_files$url[1])
```

I can download by using an index from the `ACT_MATOS_files` table above, here the file on the second row. Note that this means we have to specify what kind of data we're looking for when identifying by index.

```{r}
get_file(file = 2, project = 'umces boem offshore wind energy', data_type = 'extraction')
```

If I download using a file name, `get_file` will use the file extension to figure out what kind of data I want, so explicitly identifying the data type is not needed. Since all data extraction files are zipped, the function assumes the correct data type.
```{r}
get_file(file = 'proj87_matched_detections_2018.zip',
         project = 'umces boem offshore wind energy',
         overwrite = T)
```

## Search and download tag detections

Using the `tag_search` function, I can interface with MATOS' [tag search page](https://matos.asascience.com/search). Be very careful with this function -- it can take a *very*, **VERY** long time to return your files. This function downloads the requested CSV into your working directory, and, if `import = T` is used, reads it into your R session.

```{r, eval=FALSE}
my_detections <- tag_search(tags = paste0('A69-1601-254', seq(60, 90, 1)),
                            start_date = '03/01/2016',
                            end_date = '04/01/2016', 
                            import = T)
```

## Upload files to MATOS

There are times when you want to upload new data to MATOS. The currently accepted data types and formats are:

- newly-deployed transmitters (CSV/XLS(X))
- detection logs (CSV/VRL)
- receiver and glider deployment metadata (CSV/XLS(X))
- receiver events like temperature, tilt, etc. (CSV)
- glider GPS tracks (CSV)

A few data types use designated Ocean Tracking Network templates:

- tag metadata
- receiver deployment metadata
- glider deployment metadata

If you don't have one of these templates downloaded, you can download it through the package. For example:
```{r}
get_otn_template('glider')
```

Then, get to uploading!
```{r eval=FALSE}
post_file(project = 'umces boem offshore wind energy',
          file = c('this_is_a_dummy_file.csv', 'so_is_this.csv'),
          data_type = 'new_tags')
```


## Development

As is noted above, this package is undergoing a ton of development. If there's something I missed, please [open an issue on GitHub](https://github.com/mhpob/matos/issues) or [email me (Mike O'Brien: obrien@umces.edu) directly](mailto:obrien@umces.edu).

## Notes

- Sessions expire more-frequently than we would like. Because of this, you may often be prompted to log in. Just follow the instructions and crunch away.
