---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit this file -->
<!-- Very likely that you'll need to run rmarkdown::render('readme.rmd') rather than the knit button. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# matos

<!-- badges: start -->
<!-- badges: end -->

{matos} is a wrapper over a suite of [httr](https://httr.r-lib.org/) and [rvest](https://rvest.tidyverse.org/) functions made to interact with the [Mid-Atlantic Acoustic Telemetry Observing System website](https://matos.asascience.com/). Because of this, it's not necessarily fast or the best way to do things -- we're just pinging the website back and forth. Additionally, an internet connection is needed for pretty much anything to work.

Please note that you will need a MATOS account, [which you can sign up for here](https://matos.asascience.com/account/signup), in order to interface with any project-specific files.

## Installation

This package is still in "throw things on the wall and see what sticks" phase. You can install the most-up-to-date version from GitHub:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("mhpob/matos")
```
## Quick start

The first thing you should do is log into your MATOS account. This can be done through the package, using the [RStudio API](https://rstudio.github.io/rstudio-extensions/rstudioapi.html). This probably won't work if you're not using RStudio, so it will be changed in the future. A pop up will appear asking for your username and password. If everything works out, your credentials will be kept in the sessions' cookies. Your username/password will not be saved -- this was done intentionally so that you don't accidentally save credentials in a public script.

```{r example}
library(matos)

matos_login()
```


In addition to the above, there are/will be a few fucntions that take no arguments -- their sole function is to ping the website and return the data to you. `matos_projects` returns the [project page](https://matos.asascience.com/project), which is useful to figure out what URLs are associated with each project. You do not need MATOS permissions in order to view this page.

```{r list_projs}
all_projects <- matos_projects()

head(all_projects)
```

Since I'm logged in, I can view the files that I've uploaded to my projects.

```{r}
files <- list_files(project = 'umces boem offshore wind energy', data_type = 'project')

head(files)
```

And, I can download any of my choosing.

```{r}
files$url[1]

get_file(url = files$url[1])
```

I can also locate, download, and unzip *Data Extraction Files*.

```{r}
ACT_MATOS_files <- list_files(project = 'umces boem offshore wind energy', data_type = 'extraction')

head(ACT_MATOS_files)
```
We can download by using an index from the table above, here the file on the second row. Note that this means we have to specify what kind of data we're looking for.
```{r}
get_file(file = 2, project = 'umces boem offshore wind energy', data_type = 'extraction')
```

I can also download using a file name. Since all data extraction files are zipped, the function assumes the correct data type.
```{r}
get_file(file = 'proj87_matched_detections_2018.zip',
         project = 'umces boem offshore wind energy',
         overwrite = T)
```

