---
params:
  project_name: NULL
  project_number: NULL
  matched: NULL
  push_log: system.file("push_log.csv", package = "otndo")
  since: NULL
format:
  html:
    self-contained: true
execute:
  echo: false
title: "`r paste('Transmitter data push summary:', params$project_name)`"
author: "Created with the [`otndo` R package](https://otndo.obrien.page) (v. `r packageVersion('otndo')`)"
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)
```


```{r}
#| label: packages
#| message: false

library(otndo)
library(data.table)
library(mapview)
library(reactable)
library(leaflet)
library(leaflegend)
library(tidyr)
library(tidyverse)
```


```{r}
#| label: extraction-files-read

matched <- fread(params$matched)
matched[, day := as.Date(datecollected)]

release <- matched[receiver == "release"]
matched <- matched[receiver != "release"]

push_log <- fread(params$push_log)
```


```{r}
#| label: n-pis
#| warning: false

pis <- project_contacts(matched, type = "tag")
```

```{r}
#| label: otn-query

projects <- unique(matched$detectedby)

otn_tables <- otn_query(projects)
```

```{r}
#| label: proper-urls

## check what the proj name comes out as... send to proper URL if ACT

# if(act_eval){
#   act <- paste0('https://matos.asascience.com/project/detail/',
#                 params$project_number)
#   otn <- paste0('https://members.oceantrack.org/OTN/project?ccode=ACT.PROJ',
#                 params$project_number)
# }else{
act <- "https://matos.asascience.com/project"
otn <- "https://members.oceantrack.org/OTN/projects"
# }
```


These are data related to matched transmitter data from the `r format(push_log$date[nrow(push_log)], '%B %Y')` OTN data push. New matched-to-animal files (["matched" files](https://members.oceantrack.org/data/otn-detection-extract-documentation-matched-to-animals#autotoc-item-autotoc-1)) were uploaded on `r format(push_log$date[nrow(push_log)], '%B %d')`. The previous data push occurred on `r format(push_log$date[nrow(push_log) - 1], '%B %d, %Y')`.


## Overall

A total of `r format(nrow(matched), big.mark = ',')` detections of `r uniqueN(matched, by  = 'tagname')` individuals have been matched via [ACT](`r act`), [FACT](https://secoora.org/fact/projects/), and/or [OTN](`r otn`). The detections span across `r uniqueN(matched, by = 'detectedby')` different projects consisting of `r length(unique(unlist(strsplit(pis$PI, ',\\s?'))))` principal investigators.

```{r}
#| label: otn-match-table
#| column: page
#| fig-cap: Table 1. Total detections and number of individual fish heard by other OTN node projects.
#| fig-cap-location: top

match_table(matched, type = "tag")
```

```{r}
#| label: otn-match-map
#| fig-cap: Figure 1. Bounding box of external projects that recorded a transmitter from this project. This does not necessarily reflect the distribution of this project's fish.

otn_sf <- match_map(otn_tables)
```

```{r}
#| label: temporal-distribution
#| fig-cap: Figure 2. Temporal distribution of detections by external project, i.e., the dates when other projects heard this project's fish.

temp_dist <- temporal_distribution(matched, "tag")

temp_dist
```

```{r}
#| label: abacus-plot
#| fig-cap: Figure 3. Abacus plot of detected transmitters, colored by OTN node projects.

matched_abacus(temp_dist$data, release)
```

```{r}
#| label: station-summary-table

station_summary <- station_table(matched, type = "tag")

# station_summary |>
#   gt() |>
#   # tab_header(title = 'Records by receivers') |>
#   # tab_footnote('statons listed N-S, W-E',
#   #               locations = cells_title()) |>
#   as_raw_html()

# reactable(station_summary,
#                      groupBy = 'PI',
#                      columns = list(
#                        Project = colDef(aggregate = 'unique'),
#                        Station = colDef(aggregate = 'count'),
#                        Detections = colDef(aggregate = 'sum'),
#                        Individuals = colDef(aggregate = 'sum')
#                      )
# )
```


```{r}
#| label: detection-map-leaflet
#| fig-cap: Figure 4. Number of detections per receiver. Points represent receiver locations, while size and color reflect the number of detections at that location.

station_spatial <- prep_station_spatial(matched, type = "tag")

mapview::mapview(station_spatial,
  zcol = "Detections",
  cex = "Individuals",
  layer.name = "Matched detections"
)
```


```{r}
#| label: tags-remaining
#| fig-cap: Figure 5. Transmitters estimated to be at liberty.

remaining_transmitters(matched, push_log, release)
```


```{r results='asis', echo=FALSE}
if (is.null(params$since)) {
  # further lines are not shown (the HTML is commented out)...
  cat("<!---")

  # ...BUT they are still evaluated, so putting this in here to make the code run.
  # It will be nonsense.
  since_date <- min(matched$datelastmodified)
}
```
## What's new?

```{r}
# \ label: new-matches
if (!is.null(params$since)) {
  since_date <- as.IDate(params$since)
}
new_matches <- matched[datelastmodified > since_date]
```

Since `r format(since_date, '%B %Y')`, `r format(nrow(matched[datelastmodified > since_date]), big.mark = ',')` detections of `r uniqueN(matched[datelastmodified > since_date], by  = 'tagname')` individuals have been updated or newly-matched.

```{r}
#| label: new-otn-match-table
#| column: page
#| fig-cap: Table 2. Total newly-added detections and number of individual fish heard by other OTN node projects.
#| fig-cap-location: top

match_table(new_matches, type = "tag")
```



```{r}
#| label: new-otn-match-map
#| fig-cap: Figure 6. Bounding box of external projects providing detections added since the last data push. This does not necessarily reflect the distribution of fish heard by the project's receivers.

new_otn_tables <- lapply(
  otn_tables,
  function(.) {
    .[grepl(
      paste(gsub(".*\\.", "", unique(new_matches$detectedby)), collapse = "|"),
      collectioncode
    )]
  }
)

match_map(new_otn_tables)
```

```{r}
#| label: new-station-summary-table
#| fig-cap: Table 3. PIs, projects, and receivers providing the newly-added detections.
#| fig-cap-location: top

station_summary <- station_table(new_matches, "tag")

reactable(station_summary)
```

```{r}
# Make another label for individual
#| label: new-detection-map-leaflet

#| fig-cap: Figure 7. Number of newly-added detections per receiver. Points represent receiver locations, where color reflects the number of detections  and size  reflects the number of fish at that location.
#| 
# I find mapview has limitations in legend and it cannot show the size values. Here I use leaflet  

df <- station_spatial %>% tidyr::extract(geometry, c('lon', 'lat'), '\\((.*), (.*)\\)', convert = TRUE) 


df %>%  as.data.frame()

summary(df )

numPal <- colorNumeric('viridis', df $Detections)
leaflet(data = df) %>%
  addTiles() %>%
  addSymbolsSize(values = ~Individuals,
                 lat = ~ lat, 
                 lng = ~lon,
                 shape = 'circle',
                 color = 'black',
                 fillColor = ~numPal(Detections),
                 opacity = .1,
                 baseSize = 1) %>%
  addLegendSize(
    values = df $Individuals,
    color = 'black',
    title = 'Individual',
    shape = 'circle',
    orientation = 'horizontal',
    opacity = .1,
    fillOpacity = 0,
    breaks = 1,
    position = 'bottomright') %>%
  addLegendNumeric(
    pal = numPal, 
    title = 'Matched Detection',
    shape = 'circle',
    values = df $Detections, 
    fillOpacity = .5,
    decreasing = TRUE,
    position = 'bottomright')






```

```{r results='asis', echo=FALSE}
if (is.null(params$since)) {
  cat("-->")
}
```
