---
title: "Data Push Summaries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Push Summaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(otndo)
```

Oh, boy; oh, boy! Here it is: one of the [three yearly data pushes](https://members.oceantrack.org/faq#autotoc-item-autotoc-6) and you're ready to see all of the new data that have been matched to other projects in OTN or one of its nodes ([ACT](https://www.theactnetwork.com/), [FACT](https://secoora.org/fact/), *et al.*).

What now? What's been updated? Who might I contact for more information? Well, `make_receiver_push_summary` and `make_tag_push_summary` are here to help.

## Getting your files
The first thing you'll want to do is gather up your OTN [matched detections](https://members.oceantrack.org/data/otn-detection-extract-documentation-matched-to-animals#autotoc-item-autotoc-1) or [detection extract](https://members.oceantrack.org/data/otn-detection-extract-documentation-matched-to-animals#autotoc-item-autotoc-2) files.

### OTN projects
We'll use data from [Trudel 2018](#references) (<https://members.oceantrack.org/data/repository/pbsm>) to show how this might work. First, we'll download the files.

```{r}
# Create a folder in your temporary directory to hold the sample files
td <- file.path(tempdir(), 'otndo_test_files')
dir.create(td)

# Download deployment metadata
download.file(
  paste0(
    'https://members.oceantrack.org/data/repository/pbsm/',
    'data-and-metadata/2018/pbsm-instrument-deployment-short-form-2018.xls'
  ),
  destfile = file.path(td, 'pbsm-instrument-deployment-short-form-2018.xls'),
  # Note "mode = 'wb' is needed to download Excel files
  mode = 'wb'
)

# Download qualified detections
download.file(
  paste0(
    'https://members.oceantrack.org/data/repository/pbsm/',
    'detection-extracts/pbsm_qualified_detections_2018.zip'
  ),
  destfile = file.path(td, 'pbsm_qualified_detections_2018.zip')
)

# Download unqualified detections
download.file(
  paste0(
    'https://members.oceantrack.org/data/repository/pbsm/',
    'detection-extracts/pbsm_unqualified_detections_2018.zip'
  ),
  destfile = file.path(td, 'pbsm_unqualified_detections_2018.zip')
)
```

Now, just note where the files are saved. This will make it easier to pass into the smmary functions later.

```{r}
qualified_otn <- file.path(td, 'pbsm_qualified_detections_2018.zip')
unqualified_otn <- file.path(td, 'pbsm_unqualified_detections_2018.zip')
deployment_otn <- file.path(td, 'pbsm-instrument-deployment-short-form-2018.xls')
```

We can do the same for matched detections
```{r}
download.file(
  paste0(
    'https://members.oceantrack.org/data/repository/pbsm/',
    'detection-extracts/pbsm_matched_detections_2018.zip'
  ),
  destfile = file.path(td, 'pbsm_matched_detections_2018.zip')
)

matched_otn <- file.path(td, 'pbsm_matched_detections_2018.zip')
```


### ACT/MATOS projects
If you're a member of ACT (your project lives in the MATOS database), you can access your files via the [`matos` package](https://matos.obrien.page). Two functions in `matos` wrap `otndo`'s `make_*_summary` functions and will automatically download the necessary files for you. See [`matos::matos_tag_summary`](https://matos.obrien.page/reference/matos_tag_summary.html) and [`matos::matos_receiver_summary`](https://matos.obrien.page/reference/matos_receiver_summary.html) for more details.


### FACT projects 
At the time of this writing, there is no streamlined way to get FACT data from [Research Workspace](https://researchworkspace.com/). Before moving on to the next steps, make sure you have the necessary files downloaded.



## Running the functions
The summary functions conduct a bit of data cleaning on the front end and then run everything through a [Quarto](https://quarto.org/) or [RMarkdown](https://rmarkdown.rstudio.com/) template report. The functions use Quarto by default, but RMarkdown will be selected if:

1.  Quarto is not installed on the computer, or
2.  the `rmd` argument is set to `TRUE`.

```{r, eval=FALSE}
# Compiles with Quarto (default)
make_receiver_push_summary(qualified = qualified_otn,
                           unqualified = unqualified_otn,
                           deployment = deployment_otn,
                           rmd = F)

# Compiles with RMarkdown
make_receiver_push_summary(qualified = qualified_otn,
                           unqualified = unqualified_otn,
                           deployment = deployment_otn,
                           rmd = T)
```

Functionality is identical for `make_tag_push_summary`:

```{r, eval=FALSE}
make_tag_push_summary(matched = matched_otn)
```

## New matches "since" a certain date
Usually we want to know what has changed since the OTN nodes crossed over and talked to each other (a "data push"). This is usually when we get within-node detections, as well. These are nominally scheduled for February, July, and October. Crossover dates are stored within `otndo`; the package is updated with new dates when a data push occurs.

You can also provide a date to the "since" argument to see a summary of all of the data that have been updated since that date.

```{r, eval=FALSE}
make_tag_push_summary(matched = matched_otn,
                      since '2018-05-01')
```


## Suggestions to improve the push summaries

I am always open to suggestions on what could be added to change to make this more useful for you. Please [open an issue on GitHub](https://github.com/mhpob/matos/issues) or [email me](mailto:mike@obrien.page) with your thoughts.

## Errors and how to fix them

    Could not determine mime type for `~\Matcheddetections_layer.fgb'
    Error: pandoc document conversion failed with error 63

This error is created by an old version of the [`mapview` package](https://r-spatial.github.io/mapview/) (pre-June 2021) and has to do with the package's switch to using a [file geodatabase](https://pro.arcgis.com/en/pro-app/latest/help/data/geodatabases/manage-file-gdb/file-geodatabases.htm) to increase plotting performance. To fix this, you have two options:

1.  Update `mapview` (suggested), or
2.  Run `mapviewOptions(fgb = FALSE)` before attempting to run `make_receiver_push_summary` or `make_tag_push_summary`. Note that this will make the report build more slowly.

## References {#references}

Trudel, Marc. "A Pilot Study to Investigate the Migration of Atlantic Salmon Post-Smolts and Their Interactions with Aquaculture in Passamaquoddy Bay, New Brunswick, Canada." Ocean Tracking Network, 2018. <https://members.oceantrack.org/project?ccode=PBSM>.
